Using Genetic Algorithms to Optimize the Analogue Method for Precipitation Prediction in the Swiss Alps

### 2017-01-11 Optimization case study - JoH/elsarticle-template.tex
Reading again the JoH paper.

%TODO: read again
%TODO: Elsevier english editing
%TODO: more hydrology
%TODO: relative improvement of the crpss...

@@

This relationship is usually established by a semi-automatic sequential procedure that has strong limitations: (i) it cannot automatically choose the pressure levels and the temporal windows for a given meteorological variable, (ii) it cannot handle dependencies between the parameters, and (iii) it cannot easily handle new degrees of freedom. In this work, a global optimization approach relying on genetic algorithms could optimize all parameters jointly and automatically, which is a breakthrough in the way the analogue method was calibrated until now. It allowed taking into account parameters inter-dependencies, and selecting objectively some parameters that were manually selected beforehand (such as the pressure levels and the temporal windows of the predictor variables).

@@

The analogue method (AM) is a downscaling technique based on the idea expressed by \citet{Lorenz1956, Lorenz1969} that similar situations in terms of atmospheric circulation are likely to lead to similar local weather \citep{Duband1970}. It uses predictor variables describing the synoptic atmospheric circulation in order to predict local-scale predictands of interest. It is often used to predict daily precipitation, either in an operational forecasting context \citep[e.g.,][]{Guilbaud1997, Bontron2005, Hamill2006, Bliefernicht2010, Marty2012, Horton2012, Horton2016a, Hamill2015, BenDaoud2016} or a climate downscaling context \citep[e.g.,][]{Radanovics2013, Chardon2014, Dayon2015, Raynaud2016b}. Other predictands are also considered, such as precipitation radar images \citep{Panziera2011,Foresti2015a}, temperature \citep{Radinovic1975, Woodcock1980, Kruizinga1983, DelleMonache2013, Caillouet2016, Raynaud2016b}, wind \citep{Gordon1987, DelleMonache2013, DelleMonache2011, Vanvyve2015, Alessandrini2015, Junk2015, Junk2015c}, solar power \citep{Alessandrini2015a, Bessa2015}, snow avalanches \citep{Obled1980, Bolognesi1993}, and radiation \citep{Bois1981, Raynaud2016b}.

In operational forecasting, it is used mainly by practitioners, notably hydropower companies or flood forecasting services. They need to anticipate water yields or to issue early flood warnings several days in advance. The classical forecasting chain consists in using limited area models (e.g. AROME, or COSMO) forced by global NWP (numerical weather prediction) models with a lower resolution. However, their use require very important processing capacities, and the resulting forecast still present large uncertainties and biases. Although these outputs are essential, they can be supplemented by other sources of forecasts providing useful information. In contrast to local NWP models, AMs can transform at low cost the synoptic-scale information provided by the global NWP into precipitation forecasts, by using the natural local behavior in response to synoptic-scale influence stored in the archive of observed precipitation. Running an AM approach is so quick that it can search for analogues for every day up to 10 days ahead, eventually for the different traces of an ensemble forecast and/or issued by different NWP models (e.g. NOAA-GFS or ECMWF-IFS), in a matter of minutes.

In climate studies it is used to downscale the outputs of general circulation model (GCM) or regional climate models (RCMs) simulation runs \citep{Dayon2015} or to reconstruct past weather conditions \citep{Caillouet2016}. In future climate studies, RCMs are often used to dynamically downscale precipitation to a local scale. However, even though the relevance of RCMs'outputs increases, a bias correction of the outputs is still often required, particularly in complex terrain. Moreover, their application is computer intensive, which makes it difficult to cover all combinations of climate scenarios and GCMs. Therefore, the idea is to rely on past observations to bypass the small scale simulations and to go from the large scale situation to the end variables like precipitation by searching analogues in the archives.

@@

The study area is the alpine upper Rh\^{o}ne catchment in Switzerland (Fig.\ \ref{fig:map}). The altitude ranges from 372 to 4634~m.a.s.l.\ and the area is 5524~km$^{2}$. This region is the target of the MINERVE (Mod\'{e}lisation des Intemp\'{e}ries de Nature Extr\^{e}me sur les Rivi\`{e}res Valaisannes et de leurs Effets) project that aimed at real-time flood management on the upper Rh\^{o}ne catchment \citep{GarciaHernandez2009b}. Even though the region is rather small, the meteorological influences related to extreme weather conditions vary substantially within it \citep[see][]{Horton2012}. Indeed, a high spatial variability of the precipitation climatology exists, which is due to the complex orography of the region, and the mix of various meteorological influences. Based on different climatological analyses, the precipitation gauging stations in the catchment were clustered in 10~subregions (Fig.\ \ref{fig:map}) :

@@

Predictors are generally extracted from reanalysis datasets. NCEP-NCAR reanalysis \citep[6-hourly, 17 pressure levels at a resolution of 2.5\degree, see][]{Kalnay1996} was used here, but it could have been any other reanalysis dataset.

The predictand (which is to be predicted) is here the daily precipitation (6~a.m. to 6~a.m. the next day) measured at the MeteoSwiss' stations network, for the period 1961-2008. The time series from every available gauging station were averaged over the ten subregions of approximately 500~km$^{2}$ each in order to smooth local effects \citep{Obled2002, Marty2012}. It helps accounting for local variability, mainly when convective processes are involved, which increases slightly the prediction skill.

It must be stressed that the predictand here is a temporally cumulated variable, compared to the meteorological predictors, which may be considered instantaneous. Depending on the duration of the accumulation period (here 24h, but could have been 6, 12, 24h, or more), the choice of predictors will vary.

The 48 years precipitation dataset (see section \ref{sec:data}) was divided into a calibration period (CP) and a validation period (VP). Using data independent from the CP to validate the results is very important in order to assess the robustness of the proposed improvements and to avoid over-parametrization of the method.

In order to reduce potential bias related to trends linked to climate change or to the evolution in measurement techniques, the selection of the VP was evenly distributed over the entire series \citep{BenDaoud2010}. Thus, one year every six years were selected for validation, which represents a total of 8 years for the VP and 40 for CP. The choice of the sequence was made in order to have similar statistical characteristics between the CP and the VP.

@@

Multiple variations of the methods exist, and most of them will not be detailed hereafter \cite[see][for more comprehensive listings]{BenDaoud2016}. However, there are mainly two parameterizations that are most often used for precipitation downscaling and that were considered as reference: one that relies on an analogy of the atmospheric circulation, and another that adds a second level of analogy on moisture variables \citep{Obled2002, Bontron2005, Marty2012}.

@@

\citet{Bontron2005} show that the geopotential height at 500~hPa (Z500) and 1000~hPa (Z1000) are the best first predictors of the NCEP/NCAR reanalysis I dataset, and that the S1 criterion performs better than scores based on absolute distances. The reason for such better results is that the S1 criterion allows comparing the circulation pattern, by means of the gradients, rather than the absolute value of the geopotential height, which represent better the flow direction. To cope with seasonal effects, candidate dates are extracted within a period of 4 months centered around the target date, for every year of the archive. This method using two geopotential heights will be named 2Z.

@@

The other most known parametrization adds a second level of analogy on moisture variables (method 2Z-2MI, Table \ref{table:params_R2}). The predictor that \citet{Bontron2004} found optimal for the France territory is a moisture index made of the product of the precipitable water with the relative humidity at 850~hPa (RH850). \cite{Horton2012a} confirms that this index is also better for the Swiss Alps than any other variable from the NCEP/NCAR reanalysis I considered independently. When adding a second level of analogy, $N_{2}$ dates are subsampled within the $N_{1}$ analogues on the atmospheric circulation, to end up with a smaller number of analogue situations. When this second level of analogy is added, a higher number of analogues $N_{1}$ is kept on the first level. Moisture fields are not as well predicted by NWP models as pressure variables. It means that the 2Z-2MI method, when used in real-time forecasting, is very dependent on the skill of the NWP model in predicting moisture fields, and thus its use is often restricted to the first lead times.

@@

In order to compare the value of the score in regard to a reference, one often considers its skill score expression, and use the climatological distribution of precipitation from the entire archive as the reference. The CRPSS (\textit{Continuous Ranked Probability Skill Score}) is thus defined as following:

@@

Note however that the choice of the reference does not matter so much, since we shall consider more its relative increase or decrease rather than the CRPSS absolute value in order to assess a possible improvement.

@@

AMs are usually calibrated by a semi-automatic sequential procedure elaborated by \citet{Bontron2004} \cite[see also ][]{Radanovics2013, BenDaoud2016}. The calibration technique optimizes the spatial windows on which the predictors are compared and the number of analogues, for every level of analogy, by maximizing the performance score. However, the different analogy levels are calibrated sequentially, and the meteorological variables, the pressure levels and the temporal windows are chosen manually. The procedure, as defined by \citet{Bontron2004}, consists of the following steps:

@@

This procedure was used to calibrate the methods that were considered as references to further assess the ability of genetic algorithms to outperform the classic approach.

@@

Genetic algorithms (GAs) were developed by \citet{Holland1992b} and \citet{Goldberg1989}. They are part of Evolutionary Algorithms \citep{Back1993b, Schwefel1993}, which get inspiration from some mechanisms of biological evolution, such as reproduction, genetic mutations, chromosomal crossovers, and natural selection. GAs seek the global optimum on a complex surface, theoretically without restriction, which is of interest for AMs, that are characterized by a complex high-dimensional error function having multiple local optima. Practically, GAs allow rapidly approaching satisfactory solutions, but they do not provide the optimum solution for sure \citep{Zitzler2004a}. It is indeed mainly a matter of time. When the optimizer gets closer to the global optimum, any new improvement takes more time to appear (see for example the slow-down of the improvements over generations in Figure \ref{fig:evolution}), and the final adjustment of the parameters can be very time consuming \citep{Back1993a}. For problems that require a significant amount of time in order to evaluate the objective function, as in the case of AMs, the number of generations has to be limited in order to get reasonable processing time. Thus, different acceptable solutions can result from one or more optimization runs \citep{Holland1992b}. This is both a strength and a weakness of GAs: they are very good at exploring complex parameter spaces in order to identify the most promising areas, but they will not necessarily always find the best solution with the optimal values of all parameters \citep{Holland1992b}.

@@

The analogy of the atmospheric circulation was optimized for the 10 subregions  independently. We started from the most simple AM, and increased the complexity in order to identify the degrees of freedom that are of particular interest. Thus, the tested parametrization evolved iteratively in complexity. The detailed results of the intermediate stages are not provided in this paper \citep[see][for the details]{Horton2012a}.

The reference method for the analogy of the atmospheric circulation (2Z, Table \ref{table:params_R1}), based on Z500 and Z1000, was first considered. The optimizer had to choose simulatneously the number of analogues, both spatial windows with no overlapping constraint (i.e. they can differ from one pressure level to another), as well as the temporal windows (hours of observation of the geopotential), which cannot be achieved with the sequential calibration technique. With these degrees of freedom, a relative CRPSS improvement of 3.97\% and 2.45\% in average was obtained for the CP and the VP respectively, with regard to the reference (sequential) method based on Z500 and Z1000. Some tests showed that most of the gains were due to the non-overlapping spatial windows. This may not be seen as a large improvement, but it demonstrated that the optimizer was able to get relevant parameters.

Then, an additional degree of freedom was provided to the GAs by letting them choose the pressure levels along with the other parameters (analogue numbers, spatial and temporal windows), which is also a non-automated process in the sequential calibration. This degree of freedom increased the optimization time and may decrease the number of simulations that converge to a single solution. However, most solutions were very close in terms of score. The averaged relative improvement of the CRPSS is now 5.63\% for the CP and 3.82\% for the VP. The pressure levels that were chosen are Z500 or Z700 for the upper level, and Z925 or Z1000 (most often) for the lower level.

Parallel analyses showed that the analogy of circulation is incomplete, and that geopotential heights still contain relevant information that can improve the statistical relationship. Therefore, a third followed by a fourth predictor were added (still on the geopotential height) that the optimizer could use along with the previous parameters. There was no constraint on the predictors, so that the same pressure level could be selected more than once. Some improvements were found on the score, both for the CP and the VP, confirming that this additional information is beneficial for the quality of the prediction.

@@

The number of predictor was then successively increased up to ten, considering the weighting of the analogy criteria values. The addition of predictors globally improved the prediction skill (for both the CP and the VP) only up to four predictors (Figure \ref{fig:figure_nb_levels}). Afterwards, the score on the VP was more variable, showing eventually even a decrease, which revealed an over-parametrization of the method, and thus a lack of robustness. After four predictors, the score on the CP did not increase substantially, and could even present some local decrease due to increasing difficulty for the optimizer to converge. Selecting four predictors was considered optimal for this case study, since the gain in CRPSS was significant, and the model remained relatively simple. It cannot be excluded that another number would prevail in another region than the upper Rh\^{o}ne catchment, under other meteorological conditions, or with another reanalysis dataset.

@@

The chosen method for the atmospheric circulation analogy, based on four levels of the geopotential, and that will be named 4Zo, was made of the following degrees of freedom:

\begin{itemize}
	\setlength\itemsep{-4px}
	\item the selection of the pressure levels (4 degrees)
	\item the temporal windows (4 degrees)
	\item the spatial windows (4x4 degrees)
	\item the weights (4 degrees)
	\item the number of analogues (1 degree).
\end{itemize}

@@

The resulting CRPSS scores are provided in Figure \ref{fig:figure_crpss_4Zo} and were in average 35.8~\% for the CP and 35.5~\% for the VP, comparatively to 31.1~\% and 32.3~\% respectively for the reference method on the atmospheric circulation (optimized by the sequential procedure). The score was also calculated for three precipitation thresholds: P\(\geq\)1~mm, P\(\geq\)0.1\(\cdot\)P10 and P\(\geq\)0.5\(\cdot\)P10, P10 being the daily precipitation with a 10~year return period (Table \ref{table:scores_thresholds_4Zo}). The gain in score increased with the precipitation threshold: the relative improvement of the CRPSS was in average, respectively for the different thresholds, 13.3~\%, 15.4~\% and 29.1~\% for the CP and 7.9~\%, 11.~1\% and 34.5~\% for the VP. The optimization improved thus even more the prediction for days with significant precipitation than usual days.

@@

The resulting spatial windows (Figure \ref{fig:spatial_windows_4Zo}) may look very diverse first, but there are significant similarities for subregions located within the same vicinity. The first four subregions were characterized by a large spatial window on the upper level, whereas it was smaller elsewhere. For most subregions, the second level (Z700) was represented by thin and longitudinally extended domains. The third level (Z1000 at 6 or 12~h) also contained longitudinally extended domains, but a bit larger. The last one (Z1000 at 30~h) was rather large and squared. Subregions number 5 (Southern valleys) and 6 (Southern ridges) had exactly the same spatial windows, which suggest that they behave in a similar way and thus could have been merged. This similarity is a good sign for the accuracy of the optimized parameters.

The scores showed non-negligible improvements for both the CP and the VP (Figure \ref{fig:figure_crpss_4Zo}) compared to the reference method optimized by the sequential procedure. Even more interestingly, the results for higher precipitation thresholds (Table \ref{table:scores_thresholds_4Zo}) showed the largest improvements. This is of particular interest in the framework of flood forecasting. The further improvement of days with higher precipitation totals is likely related to the fact that larger values contribute more to the CRPS score, which means that better predicting these days results in significant increase of the performance score.

The analysis of the parameters cross-compatibility showed that obviously, the parameters were optimal on the CP when they were optimized for a given subregion (Figure \ref{fig:crossing_4Zo} top). However, the losses in CRPSS when exchanging the parameters were not of the same magnitude between the different subregions. Indeed, the Upper Rhone Valley (7) and moreover the Southeast ridges (8) seemed to behave significantly differently. These two regions have different climate properties than the others as they are particularly sensitive to southerly flows. Indeed, almost all heavy precipitation events occurred under a southerly regime, such as the Liguria, Piedmont and Aosta regions in Italy, whereas the other subregions of the catchment had heavy precipitation events mainly under a westerly regime \citep{Horton2012}. Thus, as the performance score is significantly influenced by heavy precipitation values, the parameters for the different subregions are optimized to better predict these days. It can then be expected that the optimal parameters differ between these two subregions and the other ones. This points at the importance of taking into account leading meteorological influences during discretization, that are not always best represented by the geographical distance.

Globally, the same pattern could be observed for the VP (Figure \ref{fig:crossing_4Zo} bottom), but in this case, minor improvements could occur when crossing the parameters, due to the presence of other events in the VP that might have been better predicted by a different parameter set. The relatively small gaps in score between parameterizations indicated that even though the parameters might differ significantly, the performance might not be drastically affected. Even a change in the pressure level did not mean a radical drop of the score value. A different parametrization may lead to a distinct selection of analogue days, and thus to an improvement of the prediction under certain weather conditions at the expense of others.

@@

Two methods were assessed: one with 2 moisture predictors (moisture index on 2 pressure levels or at 2 different hours) named 4Zo-2MIo, and one with four moisture predictors named 4Zo-4MIo. When introducing two predictors for the moisture analogy, the number of degrees of freedom raised to 42, and to 54 with four predictors.

@@

When optimizing a method made of 2 levels of analogy, the introduction of moisture variables in the second level had an influence on the parameter values of the first level. It means that the two levels of analogy bring complementary information, and are thus not independent. This is first visible on the number $N_{1}$ of analogues to be selected on the first level, and on the selection of the pressure levels for the circulation analogy.

@@

The selection of the temporal windows for the atmospheric circulation was similar to the preceding optimization (in the order of increasing pressure: 12~h, 24/30~h, 12~h, 30~h), but sometimes with a bit more variability. When it comes to the moisture analogy, there was a clear trend to select 12~h and 24~h when considering 2 predictors. However, it must be remembered that this holds for our predictand, the accumulated precipitation over 06-30~h~UTC, and that it is expected to differ if we change the temporal window (e.g. 00-24~h~UTC) or its duration.

The optimized spatial windows for the atmospheric circulation have also changed (Figure \ref{fig:spatial_windows_4Zo-2MIo}; results of 4Zo-4MIo are not shown but present the same trends). The very large domains on the upper level of the four first subregions were not present anymore, and more variability could be observed. The selected points for the analogy on the moisture index were always located nearby the catchment, including at least 1 of the nearest points from the reanalysis dataset, and the spatial windows were relatively small. Thus, for this case study, there is no need to look for distant moisture information and the search could be reduced to a smaller domain.

>>>>

%TODO: read again
%TODO: Elsevier english editing
%TODO: relative improvement of the crpss...
%TODO: homogenize hours: 6 a.m. or 06 h ??

@@

This relationship is usually established by a semi-automatic sequential procedure that has strong limitations: (i) it cannot automatically choose the pressure levels and the temporal windows for a given meteorological variable, (ii) it cannot handle dependencies between the parameters, and (iii) it cannot easily handle new degrees of freedom. In this work, a global optimization approach relying on genetic algorithms could optimize all parameters jointly and automatically. It allowed taking into account parameters inter-dependencies, and selecting objectively some parameters that were manually selected beforehand, which e.g. obviates the need of assessing a high number of combinations of the pressure levels and the temporal windows of the predictor variables.

@@

The analogue method (AM) is a downscaling technique based on the idea expressed by \citet{Lorenz1956, Lorenz1969} that similar situations in terms of the atmospheric circulation are likely to lead to similar local weather \citep{Duband1970}. It uses predictor variables describing the synoptic atmospheric circulation in order to predict local-scale predictands of interest. It is often used to predict daily precipitation, either in an operational forecasting context \citep[e.g.,][]{Guilbaud1997, Bontron2005, Hamill2006, Bliefernicht2010, Marty2012, Horton2012, Horton2016a, Hamill2015, BenDaoud2016} or a climate downscaling context \citep[e.g.,][]{Radanovics2013, Chardon2014, Dayon2015, Raynaud2016b}. Other predictands are also considered, such as precipitation radar images \citep{Panziera2011,Foresti2015a}, temperature \citep{Radinovic1975, Woodcock1980, Kruizinga1983, DelleMonache2013, Caillouet2016, Raynaud2016b}, wind \citep{Gordon1987, DelleMonache2013, DelleMonache2011, Vanvyve2015, Alessandrini2015, Junk2015, Junk2015c}, solar power \citep{Alessandrini2015a, Bessa2015}, snow avalanches \citep{Obled1980, Bolognesi1993}, and radiation \citep{Bois1981, Raynaud2016b}.

In real-time forecasting, it is used mainly by practitioners, notably hydropower companies or flood forecasting services. They need to anticipate water yields or to issue early flood warnings several days in advance. The classical forecasting chain consists in using limited area models (e.g. AROME, or COSMO) forced by global NWP (numerical weather prediction) models with a lower resolution. However, their use require very important processing capacities, and the resulting forecast still present large uncertainties and biases. Although these outputs are essential, they can be supplemented by other sources of forecasts providing useful information. In contrast to local NWP models, AMs can transform at low cost the synoptic-scale information provided by the global NWP model into precipitation forecasts, by using the natural local behavior in response to synoptic-scale influence stored in the archive of observed precipitation. Running an AM approach is so quick that it can search for analogues for every day up to ten days ahead, eventually for the different traces of an ensemble forecast and/or issued by different NWP models (e.g. NOAA-GFS or ECMWF-IFS), in a matter of minutes.

In climate studies AMs are used to downscale the outputs of general circulation model (GCM) or regional climate models (RCMs) simulation runs \citep{Dayon2015} or to reconstruct past weather conditions \citep{Caillouet2016}. In future climate studies, RCMs are often used to dynamically downscale precipitation to a local scale. However, even though the relevance of RCMs'outputs increases, a bias correction of the outputs is still often required, particularly in complex terrain. Moreover, their application is computer-intensive, which makes it difficult to cover all combinations of climate scenarios and GCMs. Therefore, the idea is to bypass the small scale simulations and to go from the large scale situation to the end variables like precipitation by statistical downscaling \citep{Maraun2010}.

@@

The study area is the alpine upper Rh\^{o}ne catchment in Switzerland (Fig.\ \ref{fig:map}). The altitude ranges from 372 to 4634~m.a.s.l.\ and the area is 5524~km$^{2}$. This region is the target of the MINERVE (Mod\'{e}lisation des Intemp\'{e}ries de Nature Extr\^{e}me sur les Rivi\`{e}res Valaisannes et de leurs Effets) project that aimed at real-time flood management on the upper Rh\^{o}ne catchment \citep{GarciaHernandez2009b}. Even though the region is rather small, the meteorological influences related to extreme weather conditions vary substantially within it \citep[see][]{Horton2012}. Indeed, a high spatial variability of the precipitation climatology exists, which is due to the complex orography of the region, and the mix of various meteorological influences. Based on different climatological analyses, the precipitation gauging stations in the catchment were clustered in ten subregions (Fig.\ \ref{fig:map}) :

@@

Predictors are generally extracted from reanalysis datasets. The NCEP-NCAR reanalysis I \citep[6-hourly, 17 pressure levels at a resolution of 2.5\degree, see][]{Kalnay1996} was used here, but it could have been any other reanalysis dataset.

The predictand (which is to be predicted) is here the daily precipitation (6~a.m. to 6~a.m. the next day) measured at the MeteoSwiss' stations network, for the period 1961-2008. The time series from every available gauging station were averaged over the ten subregions (Fig.\ \ref{fig:map}) of approximately 500~km$^{2}$ each in order to smooth local effects \citep{Obled2002, Marty2012}. It helps accounting for local variability, mainly when convective processes are involved, which increases slightly the prediction skill.

It must be stressed that the predictand is here sa temporally cumulated variable, compared to the meteorological predictors, which may be considered instantaneous. Depending on the duration of the accumulation period (here 24~h, but could have been 6, 12~h, or more than 24~h), the choice of predictors will vary.

The 48 years precipitation dataset was divided into a calibration period (CP) and a validation period (VP). Using data independent from the CP to validate the results is very important in order to assess the robustness of the proposed improvements and to avoid over-parametrization of the method.

In order to reduce potential bias related to trends linked to climate change or to the evolution in measurement techniques, the selection of the VP was evenly distributed over the entire series \citep{BenDaoud2010}. Thus, one year every six years was selected for validation, which represents a total of 8 years for the VP and 40 for CP. The choice of the sequence was made in order to have similar statistical characteristics between the CP and the VP.

@@

Multiple variations of the analogue method exist, and most of them will not be detailed hereafter \cite[see][for more comprehensive listings]{BenDaoud2016}. However, there are mainly two parameterizations that are most often used for precipitation prediction and that were considered as reference: one that relies on an analogy of the atmospheric circulation, and another that adds a second level of analogy on moisture variables \citep{Obled2002, Bontron2005, Marty2012}.

@@

\citet{Bontron2005} show that the geopotential height at 500~hPa (Z500) and 1000~hPa (Z1000) are the best first predictors of the NCEP/NCAR reanalysis I dataset, and that the S1 criterion performs better than scores based on absolute distances. The reason for such better results is that the S1 criterion allows comparing the circulation patterns, by means of the gradients, rather than the absolute value of the geopotential height, which represent better the flow direction. To cope with seasonal effects, candidate dates are extracted within a period of four months centered around the target date, for every year of the archive. This method using two geopotential heights will be named 2Z.

@@

The other most known parametrization adds a second level of analogy on moisture variables (method 2Z-2MI, Table \ref{table:params_R2}). The predictor that \citet{Bontron2004} found optimal for the France territory is a moisture index made of the product of the total precipitable water (TPW) with the relative humidity at 850~hPa (RH850). \cite{Horton2012a} confirms that this index is also better for the Swiss Alps than any other variable from the NCEP/NCAR reanalysis I considered independently. When adding a second level of analogy, $N_{2}$ dates are subsampled within the $N_{1}$ analogues on the atmospheric circulation, to end up with a smaller number of analogue situations. When this second level of analogy is added, a higher number of analogues $N_{1}$ is kept on the first level. Moisture fields are not as well predicted by NWP models as pressure variables. It implies that the 2Z-2MI method, when used in real-time forecasting, is very dependent on the skill of the NWP model in predicting moisture fields, and thus its use is often restricted to the first lead times.

@@

In order to compare the value of the score in regard to a reference, one often considers its skill score expression, and use the climatological distribution of precipitation from the entire archive as the reference. The CRPSS (Continuous Ranked Probability Skill Score) is thus defined as following:

@@

Note however that the choice of the reference does not matter so much when assessing potential improvements of the method, since we shall consider more its relative increase or decrease rather than the CRPSS absolute value.

@@

AMs are usually calibrated by a semi-automatic sequential procedure elaborated by \citet{Bontron2004} \cite[see also ][]{Radanovics2013, BenDaoud2016}. The calibration technique optimizes the spatial windows on which the predictors are compared and the number of analogues, for every level of analogy, by maximizing the performance score (CRPSS). However, the different analogy levels are calibrated sequentially, and the meteorological variables, the pressure levels and the temporal windows are chosen manually. The procedure, as defined by \citet{Bontron2004}, consists of the following steps:

@@

This procedure was used to calibrate the methods that were here considered as references to further assess the ability of genetic algorithms to outperform the classic approach.

@@

Genetic algorithms (GAs) were developed by \citet{Holland1992b} and \citet{Goldberg1989}. They are part of Evolutionary Algorithms \citep{Back1993b, Schwefel1993}, which get inspiration from some mechanisms of biological evolution, such as reproduction, genetic mutations, chromosomal crossovers, and natural selection. GAs seek the global optimum on a complex surface, theoretically without restriction, which is of interest for AMs, that are characterized by a complex high-dimensional error function having multiple local optima. Practically, GAs allow rapidly approaching satisfactory solutions, but they do not provide the optimum solution for sure \citep{Zitzler2004a}. It is indeed mainly a matter of time. When the optimizer gets closer to the global optimum, any new improvement takes more time to appear (see for example the slow-down of the improvements over generations in Figure \ref{fig:evolution}), and the final adjustment of the parameters can be very time consuming \citep{Back1993a}. For problems that require a significant amount of time in order to evaluate the objective function, as in the case of AMs (because it needs to make a prediction for every day of the CP), the number of generations has to be limited in order to get reasonable processing time. Thus, different acceptable solutions can result from one or more optimization runs \citep{Holland1992b}. This is both a strength and a weakness of GAs: they are very good at exploring complex parameter spaces in order to identify the most promising areas, but they will not necessarily always find the best solution with the optimal values of all parameters \citep{Holland1992b}.

@@

The analogy of the atmospheric circulation was optimized for the ten subregions (Section \ref{sec:case_study})  independently. We started from the most simple AM, and increased the complexity in order to identify the degrees of freedom that are of particular interest. Thus, the tested parametrization evolved iteratively in complexity. The detailed results of the intermediate stages are not provided in this paper \citep[see][for the details]{Horton2012a}.

The reference method for the analogy of the atmospheric circulation (2Z, Table \ref{table:params_R1}), based on Z500 and Z1000, was first considered. The optimizer had to choose simulatneously the number of analogues, both spatial windows with no overlapping constraint (i.e. they can differ from one pressure level to another), as well as the temporal windows (hours of observation of the geopotential), which cannot be achieved with the sequential calibration technique. The performance score (CRPSS) was slightly improved, with these limited degrees of freedom, in regards to the 2Z reference method calibrated with the sequential procedure. Some tests showed that most of the gains were due to the non-overlapping spatial windows. It demonstrated that the optimizer was able to get relevant parameters for a simple method.

Then, an additional degree of freedom was provided to the GAs by letting them choose the pressure levels along with the other parameters (analogue numbers, spatial and temporal windows), which is also a non-automated process in the sequential calibration. This degree of freedom increased the optimization time and might decrease the number of simulations that converge to a single solution. However, most solutions were very close in terms of the performance score, which was improved further. The selected pressure levels were Z500 or Z700 for the upper level, and Z925 or Z1000 (most often) for the lower level.

Parallel analyses showed that the analogy of circulation is incomplete, and that geopotential heights still contain relevant information that can improve the statistical relationship. Therefore, a third, followed by a fourth predictor were added (still only geopotential heights). There was no constraint on the predictors, so that the same pressure level could be selected more than once. Further improvements were found on the performance score, both for the CP and the VP, confirming that this additional information was beneficial for the quality of the prediction.

@@

The number of predictors (still only geopotential heights) was then successively increased up to ten, considering the weighting of the analogy criteria values. The addition of predictors globally improved the prediction skill (for both the CP and the VP) only up to four predictors (Figure \ref{fig:figure_nb_levels}). Afterwards, the score on the VP was more variable, showing eventually even a decrease, which revealed an over-parametrization of the method, and thus a lack of robustness. After four predictors, the score on the CP did not increase substantially, and could even present some local decrease due to increasing difficulty for the optimizer to converge. Selecting four predictors (geopotential heights) was considered optimal for this case study, since the gain in CRPSS was significant, and the model remained relatively simple. It cannot be excluded that another number would prevail in another region than the upper Rh\^{o}ne catchment, under other meteorological conditions, or with another reanalysis dataset.

@@

The chosen method for the atmospheric circulation analogy, based on four predictors of geopotential heights, and that will be named 4Zo (o for optimized), was made of the following degrees of freedom:

\begin{itemize}
	\setlength\itemsep{-4px}
	\item the selection of the pressure levels (4 degrees),
	\item the temporal windows (4 degrees),
	\item the spatial windows (4x4 degrees),
	\item the weights (4 degrees),
	\item the number of analogues (1 degree).
\end{itemize}

@@

The resulting CRPSS scores are provided in Figure \ref{fig:figure_crpss_4Zo} and were in average 35.8~\% for the CP and 35.5~\% for the VP, comparatively to 31.1~\% and 32.3~\% respectively for the reference method 2Z on the atmospheric circulation (optimized by the sequential procedure). The score was also calculated for three precipitation thresholds: P\(\geq\)1~mm, P\(\geq\)0.1\(\cdot\)P10 and P\(\geq\)0.5\(\cdot\)P10, P10 being the daily precipitation with a 10~year return period (Table \ref{table:scores_thresholds_4Zo}). The gain in score increased with the precipitation threshold: the relative improvement of the CRPSS was in average, respectively for the different thresholds, 13.3~\%, 15.4~\% and 29.1~\% for the CP and 7.9~\%, 11.~1\% and 34.5~\% for the VP. The optimization improved thus even more the prediction for days with significant precipitation than usual days.

@@

The resulting spatial windows (Figure \ref{fig:spatial_windows_4Zo}) may look very diverse first, but there are significant similarities for subregions located within the same vicinity. The first four subregions were characterized by a large spatial window on the upper level, whereas it was smaller elsewhere. For most subregions, the second level (Z700) was compared on thin and longitudinally extended spatial windows. The third level (Z1000 at 6 or 12~h) also had longitudinally extended domains, but a bit larger. The last one (Z1000 at 30~h) had rather large and squared windows. Subregions number 5 (Southern valleys) and 6 (Southern ridges) had exactly the same spatial windows, which suggest that they behave in a similar way and thus could have been merged. This similarity is a good sign for the accuracy of the optimized parameters.

The performance scores showed non-negligible improvements for both the CP and the VP (Figure \ref{fig:figure_crpss_4Zo}) compared to the 2Z reference method optimized by the sequential procedure. Even more interestingly, the results for higher precipitation thresholds (Table \ref{table:scores_thresholds_4Zo}) showed the largest improvements. This is of particular interest in the framework of flood forecasting. The further improvement of days with higher precipitation totals is likely related to the fact that larger values contribute more to the CRPS score, which means that better predicting these days results in significant increase of the global performance score.

The analysis of the parameters cross-compatibility showed that obviously the parameters were optimal on the CP for the subregion for which they were optimized (Figure \ref{fig:crossing_4Zo} top). However, the losses in CRPSS when exchanging the parameters were not of the same magnitude between the different subregions. Indeed, the Upper Rhone Valley (7) and moreover the Southeast ridges (8) seemed to behave significantly differently. These two regions have different climatic properties than the others as they are particularly sensitive to southerly flows. Indeed, almost all heavy precipitation events occurred under a southerly regime, such as in the Liguria, Piedmont and Aosta regions in Italy, whereas the other subregions of the catchment had extreme events mainly under a westerly regime \citep{Horton2012}. Thus, as the performance score is significantly influenced by heavy precipitation values, the parameters for the different subregions are likely optimized to better predict these days. It can then be expected that the optimal parameters differ between these two subregions and the other ones. This points at the importance of taking into account leading meteorological influences during precipitation stations clustering, that are not always best represented by the geographical distance.

Globally, the same pattern could be observed for the VP (Figure \ref{fig:crossing_4Zo} bottom), but in this case, minor improvements could occur when crossing the parameters, due to the presence of other events in the VP that might have been better predicted by a different parameter set. The relatively small differences in score between parameterizations indicated that even though the parameters might differ significantly, the performance might not be drastically affected. Even a change in the pressure level did not mean a radical drop of the score value. A different parametrization may lead to a distinct selection of analogue days, and thus to an improvement of the prediction under certain weather conditions at the expense of others.

@@

Two methods were assessed: one with two moisture predictors (moisture index on two pressure levels or at two different hours) named 4Zo-2MIo, and one with four moisture predictors named 4Zo-4MIo. When introducing two predictors for the moisture analogy, the number of degrees of freedom raised to 42, and to 54 with four predictors.

@@

When optimizing a method made of two levels of analogy, the introduction of moisture variables in the second level had an influence on the parameter values of the first level. It means that the two levels of analogy bring complementary information, and are thus not independent. This is first visible on the number $N_{1}$ of analogues to be selected on the first level, and on the selection of the pressure levels for the circulation analogy.

@@

The selection of the temporal windows for the atmospheric circulation was similar to the preceding optimization (in the order of increasing pressure: 12~h, 24/30~h, 12~h, 30~h), but sometimes with a bit more variability. When it comes to the moisture analogy, there was a clear trend to select 12~h and 24~h when considering two predictors. However, it must be remembered that this holds for our predictand, the accumulated precipitation over 06-30~h~UTC, and that it is expected to differ if we change the temporal window (e.g. 00-24~h~UTC) or its duration.

The optimized spatial windows for the atmospheric circulation have also changed (Figure \ref{fig:spatial_windows_4Zo-2MIo}; results of 4Zo-4MIo are not shown but present the same trends). The very large domains on the upper level of the four first subregions were not present anymore, and more variability could be observed. The selected points for the analogy on the moisture index were always located nearby the catchment, including at least one of the nearest points from the reanalysis dataset, and the spatial windows were relatively small. Thus, for this case study, there is no need to look for distant moisture information and the search could be reduced to a smaller domain.
